---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(omxr)
```
# omxr
[![CRAN status](https://www.r-pkg.org/badges/version/omxr)](https://cran.r-project.org/package=omxr)
[![Build Status](https://travis-ci.org/gregmacfarlane/omxr.svg?branch=master)](https://travis-ci.org/gregmacfarlane/omxr)


The goal of omxr is to provide a convenient API for R users working with 
Open Matrix (OMX) format files. For more details on the file standard, see
the [osPlanning wiki](https://github.com/osPlanning/omx/wiki).

## Installation

We hope to have omxr on [CRAN](https://CRAN.R-project.org) in the near future. 
Until then, you can install omxr with the [devtools](https://cran.r-project.org/package=devtools) package

``` r
devtools::install_github("gregmacfarlane/omxr")
```

## Example

With omxr you can examine the contents of existing OMX files within R.
```{r example}
# path to the omx file
omx_file <- system.file("extdata", "test.omx", package = "omxr")

# list contents
list_omx(omx_file)
```

You can choose to read the matrix as either an R matrix or as a tidyverse-friendly 
tibble, depending on your needs.

```{r read}
m <- read_omx(omx_file, "test") # return an array
m[1:5, 1:5]

read_all_omx(omx_file)
```

More detailed examples, including creating and writing to OMX files, are
available in the package vignette.

``` r
vignette(omxr)
```

