---
title: "Using `omxr`"
author: "Greg Macfarlane"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_caption: TRUE
vignette: >
  %\VignetteIndexEntry{omxr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup}
set.seed(10)
```



```{r load_library}
library(omxr)
```


# Create and write.

`.omx` files are HDF5 containers storing many different matrices and their
attributes. You can create a new container with the `create_omx` command.

```{r create_omx}
zones <- 1:10
create_omx( file = "simple.omx", numrows = length(zones), numcols = length(zones))
```

Let's start with a 10-zone system with random trips and costs.

```{r make_matrix}
trips <- matrix(rnorm(n = length(zones)^2, 200, 50),  
                nrow = length(zones), ncol = length(zones))
cost <- matrix(rlnorm(n = length(zones)^2, 1, 1),
               nrow = length(zones), ncol = length(zones))
```

We can store these in the omx container.

```{r write_omx}
write_omx(file = "simple.omx", matrix = trips, "trips", 
          description = "Total Trips")

write_omx(file = "simple.omx", matrix = cost, "cost", 
          description = "Generalized Cost")
```


# Read

Now that the matrices are in the file, we can try reading them.

```{r read_matrix}
read_omx("simple.omx", "trips")
read_omx("simple.omx", "cost")
```

You can transform a matrix to long format, which is more `tidyverse`
friendly.

```{r long_matrix(), message=FALSE, warning=FALSE}
library(tidyverse)
read_omx("simple.omx", "trips") %>%
  long_matrix(value = "trips")
```

You can also read subsets of matrices.

```{r read_subset}
read_omx("simple.omx", "trips", row_index = 2:4, col_index = 2:5)
```

There are also functions to get the attributes of an OMX file. A call to
`list_omx()` will show the names of the matrices inside the file.

```{r attributes}
get_omx_attr("simple.omx")
list_omx("simple.omx")
```


# Lookups

Perhaps a better way to read a subset of a matrix is with a defined "lookup." 
These can be stored in the omx container in addition to the matrices. You specify
the lookup with a vector of the appropriate length, where `TRUE` means the 
row or column should be included, and `FALSE` means otherwise.

```{r write_lookup}
lookup <- zones %in% c(1, 2:5, 9)
lookup
write_lookup("simple.omx", lookup_v = lookup, 
             name = "trial", description = "test lookup", replace = TRUE)
```

Now we can get the selected rows and columns from either matrix automatically.

```{r read_lookup}
read_selected_omx("simple.omx", "trips", 
                  row_selection = "trial", col_selection = "trial")
```

